Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data

### 2021-01-16 documents/ovando-etal-assessing-global-fisheries.Rmd
submitted to proc b

---
title: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: proceedings-of-the-royal-society-b.csl
output: 
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fish populations play an important role in tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain. Numerous high-profile publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added. Obtaining accurate estimates of stock status for the world's fisheries depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
---

@@

data limited are not just messier versions of data rich

80% in trouble

@@

The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development".While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020]. Meeting the SDG 14 targets depends in part on improving our ability to effectively track the status of global marine fish populations.

The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's report on the State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries. The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020]. Where possible, statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database [@ricard2012]. The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the biological state of "assessed" fish populations around the globe.

That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed". The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020]. While these methods combined with local knowledge can be well informed as to the status of their fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods. With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, and modeling methods,the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.

In this paper we evaluate the ability of alternative data-limited assessment models to improve our estimate of the population status of unassessed fish stocks around the world. While these unassessed stocks are generally individually smaller than the typically larger and more valuable fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts use forms of "catch-only" stock assessment models (Free et al. 2020 [@free2020] and references therein). These models seek to infer stock status, for example biomass B relative to the biomass at maximum sustainable yield B~MSY~, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013].

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates B/B~MSY~ and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable methods intended for use when insufficient data are available for a full stock assessment model.

Costello et al. 2016 [@costello2016a] finds similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in Rosenberg et al. 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.

@@

readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
  select(`FAO Area`,`FAO % Overfished`, everything()) %>% 
  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

## Experiment Structure

We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global fish populations. We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history. B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data. This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.

We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM. We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al. 2017 [@froese2017] (See Table.\@ref(tab:dat-desc)). We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).

@@

We allow for process error *p* (in the manner of the stochastic stock reduction analysis error suggested by [@walters2006]). This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average. Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).

@@

All of our estimates are Bayesian in nature. sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent). Unless there is an abundance index to fit to the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis. All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.

@@

Suppose that the only thing we observe from a fishery is a catch history. Assuming Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past. We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history. Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero.

In the absence of any data to fit to, the SRA algorithm works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief. This creates a problem for the Bayesian nature of our analysis though. Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*. Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step. Doing so places essentially two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context). This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood. However, particularly for the SRA version of sraplus, Borel's Paradox poses a particular problem.

The SRA algorithm works in two steps. First, the algorithm rejects any draws that resulted in the collapse of the population (biomass less than catch in a given timestep). From there a standard SRA would sample from the priors in proportion to the stated prior on recent stock status. If the bulk of the prior on terminal stock status was concentrated at 50% of *K*, combinations of *r* and *K* that produce terminal stock status near 50% of *K* are sampled proportionally more frequently. However, lower values of terminal stock status have fewer candidate values of *r* and *K*, since it becomes harder and harder to find viable pairs that come close to but do not crash the population at any time step. Conversely, in the absence of constraints higher values of stock status have infinite combinations of plausible *r* and *K* combinations: since under this model the population cannot be greater than carrying capacity, as for example *K* approaches infinity terminal stock status asymptotes at close to 100% of *K*. The net result of this is that even though individual combinations of *r* and *K* that produce higher stock status than the mean of the prior on recent stock status individually have lower probability of being sampled, there are many more opportunities for the lower-probability events that produce higher stock status to be sampled. As a result, the post-model-pre-data prior on terminal depletion will always be higher under this method than the supplied prior on stock status.

We use an approximation to this problem here, similar in spirit to Bayesian melding [@poole2000]. Our solution amounts to a two-step SIR algorithm. We first run the standard SIR algorithm as described above. We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin.

$$p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}$$

We then divide the sampling probability of of bin *i* evenly among each of the draws within that bin *n*

$$p(n_i) = \frac{p(bin_i)}{N_i}$$

And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$. The net result of this is a post-model-pre-data distribution of parameters *r* and *K* that produce a distribution of recent stock status that roughly matches the supplied prior on recent stock status. In effect, this process answers the question "given the model, what combinations of parameters produce my prior on recent stock status". This is only an approximate solution, but it helps ensure that the post-model-pre-data distribution of stock status much more closely matches the stated prior on recent stock status, and reduced the positive bias resulting from use of the raw SRA algorithm (Fig.\@ref(fig:bias-plot), Fig.S1).

```{r bias-plot, fig.cap = "Post-model-pre-data distribution of depletion (biomass relative to carrying capacity) from raw SRA algorithm (untuned, top row), from SRA algorithm with approximate tuning applied (tuned, middle row), compared to the supplised prior on depletion (bottom row). Black vertical line indicates median value. "}

id = unique(ram_comp_data$stockid)

ex <- ram_data %>% 
  filter(stockid %in% id[[1]])

# fit using catch heuristic

co_driors <- sraplus::format_driors(taxa = unique(ex$scientificname),
                                 catch = ex$catch, 
                                 years = ex$year,
                                 terminal_state = 0.25,
                                 terminal_state_cv = 0.5,
                                 b_ref_type = "k")

biased_co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = FALSE,
                                      estimate_shape = FALSE, estimate_proc_error = TRUE)

biased_depletion <- biased_co_fit$fit %>% 
  filter(variable == "dep_t") %>%
  filter(year == max(year)) %>% 
  mutate(set = "Untuned Post-Model-Pre-Data") %>% 
  select(year, value, set)

co_fit <- sraplus::fit_sraplus(driors = co_driors, tune_prior_predictive = TRUE,
                                      estimate_shape = FALSE, 
                               estimate_proc_error = TRUE)

tuned_depletion <- co_fit$fit %>% 
  filter(year == max(year), variable == "dep_t") %>% 
  mutate(set = "Tuned Post-Model-Pre-Data") %>% 
  select(year, value, set)

prior <- tuned_depletion %>% 
  mutate(value = rlnorm(n(), log(0.25), 0.5)) %>% 
    mutate(set = "Prior")

comps <- biased_depletion %>% 
  bind_rows(tuned_depletion) %>% 
  bind_rows(prior) %>% 
  group_by(set) %>% 
  mutate(valrank = percent_rank(value)) %>% 
  filter(between(valrank, .05, .95))

comps %>% 
  ggplot(aes(x = value, y = set)) + 
  ggridges::geom_density_ridges(alpha = 0.9,
                                quantile_lines = TRUE, quantiles = 2) +
  scale_x_continuous(limits = c(0, 1.2), name = "Depletion (B/K)") + 
  scale_y_discrete(name = '')

@@

Along with allowing users to supply their own priors, the sraplus package contains three built-in methods used in this manuscript for converting information on stock status from additional outside data into a form usable as a prior by sraplus. We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database. We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types. Given values of these variables for a new fishery then, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus.

All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017]. The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus. Given the scope of this analysis, we do not claim that the presented regressions are the best possible model relating these data with the fishery status indicators of interest. Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.

#### Catch-Only Priors {.unnumbered}

Many of the current methods for estimating global stock status of unassessed stocks are based on predicting stock status from characteristics of the catch history [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018]. While these catch-only methods have been shown to have serious shortcomings [@free2020], we include them as a point of reference given their ubiquity in the global assessment literature.

We used data from the RAM Legacy Stock Assessment Database to estimate a regression of stock status as a function of catch history characteristics. To facilitate the process, we first fit a spectral clustering algorithm to the scaled catch histories of fisheries in RAM, in order identify four possible clusters of catch history types within the the data. Emergent clusters show for example one built around a downward "one way trip" style catch histories, others with a boom and bust pattern, others with stable but fluctuating catches.

We then trained a classification algorithm to predict which catch cluster a given fishery would fall into based on the shape of its catch history. This algorithm was then used to assign fisheries to one of the four identified catch history types, and the catch history type was then used as a hierarchical term within our catch-based regressions (where *s* refers to a smoothing term). For the first regression, we restrict the data to the first year of data available for each fishery *i*, in order to estimate initial stock status

$$log(value_i) \sim normal(s(\frac{first(catch)}{max(catch)} | cluster_i) + s(log(lengthi)|cluster_i) + 1, \sigma)$$

For the second regression, we included data for all available years *y* for fishery *i*. The model is then used to construct a prior on fishery status in the terminal year of the data

$$log(value_{i,y}) \sim normal(s(fyear | cluster_i) + s(\frac{catch_{i,y}}{max(catch_i)} | cluster_i) + cluster_i, \sigma)$$

where *fyear* is the year of the fishery, starting from 0.

#### Fisheries Management Index Priors {.unnumbered}

The Fisheries Management Index (FMI), as presented in [@melnychuk2017], utilizes surveys filled out by regional experts to score a fishery against a set of 46 specific questions for individual species about what elements of fisheries management were in place. These questions are then aggregated into broader categories of science, enforcement, management, and socioeconomic. The higher the score, the better the expert judges that a given metric is met in that fishery. Importantly, FMI surveys can be filled out in the absence of stock assessments. This allows us to explore how FMI values map onto stock status, and explore the ability then to use FMI scores to produce priors on stock status for unassessed fisheries (in a manner similar to [@osio2015] and [@cope2015]).

The final selected model relating FMI variable to stock status metrics was a generalized additive model (GAM) of the form

$$log(value_i) \sim N(s(research_i) + s(management_i) +
s(enforcement_i) + s(socioeconomics_i) + \frac{catch_i}{max(catch)_i)} + 1,\sigma_{SAR})$$

#### Swept Area Ratio Priors {.unnumbered}

[@amoroso2018] provides an extensive database of trawling footprints throughout the world, including both regions heavily covered by stock assessments and largely unassessed areas. This makes the trawl footprint data an ideal candidate for supporting global stock assessment efforts. As illustrated in [@amoroso2018], there is an evident positive relationship between the swept area ratio (SAR,the total annual area trawled divided by the total area of the region) and U/U~MSY~. Note that SAR can be greater than 1 since the same area can be trawled multiple times in a year, e.g. if all trawl-able areas are trawled twice a year then the SAR will be 2. Also note the skewed distribution of SAR values with most concentrated well below 1 and only a handful above 1.

The final selected model relating SAR to to stock status metrics was

$$log(value_i) \sim normal(s(SAR_i) + s(\frac{catch_i}{max(catch)_i)}) + 1,\sigma_{SAR}) $$

@@

We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries. The VOI analysis was performed by using sraplus to generate estimates of stock status (B/B~MSY~) for stocks in the RAM legacy stock assessment, and comparing the estimated values to the values reported in RAM. We generate fits for 3000 combinations of a RAM stock and available data. For any one draw, we randomly sample a RAM stock and a list of available data and data quality. For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance. The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.

Using this set of fits, we assess performance as the root-mean-squared-error of B/B~MSY~ over the most recent 5 years of the fishery, in order to evaluate the ability of the model to capture the recent trends in stock status and not just the most recent year. We evaluate the contributing of each data type to RMSE using a Gamma GLM with a log link of the form

$$rmse \sim  Gamma(\pmb{\beta}{\pmb{X}} + (1 | stock), shape, scale)$$

Where $\pmb{\beta}$ is the vector of coefficients associated with the matrix of dummy variables marking the use of different data types in the vector $\pmb{X}$

@@

We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index (FMI) scores (a measure of management capacity), and swept area ratio (SAR) values (a measure of trawl fishing intensity) (Table.\@ref(tab:data)).

@@

# Supplementary Information {.unnumbered}

\renewcommand{\thefigure}{S\arabic{figure}}

\setcounter{figure}{0}

>>>>

---
title: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: proceedings-of-the-royal-society-b.csl
output: 
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fish populations play an important role in tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain. Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added. Estimates of fish populations based primarily on catch histories alone frequently performed roughly as well as a random guess. Obtaining accurate estimates of stock status for the world's exploited fish populations depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
---

@@

<!-- data limited are not just messier versions of data rich  -->

<!-- 80% in trouble  -->

@@

The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". Marine fisheries represent one of the largest anthropogenic impacts on the oceans, and when managed sustainably are a critical source of economic prosperity and food security around the globe. As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries. While our understanding of the state of world fisheries has improved over the last decade the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their biological status [@hilborn2020;@fao2020]. This is a major impediment to ensuring the sustainable development of the world's oceans. In this paper, we assess why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.

The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries. The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020]. Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012]. The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the biological state of assessed fish populations around the globe.

That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed". The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020]. While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods. With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals.

We evaluate the ability of alternative assessment models to improve our estimate of the population status of unassessed fish stocks around the world. While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of unassessed fish stocks lacking the data needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts use forms of "catch-only" data-limited models (Free et al. 2020 [@free2020] and references therein). These models seek to infer stock status, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries (e.g. @cabral2020,@gaines2018).

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*. These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by potentially less reliable catch-only models intended for use when insufficient data are available for a full stock assessment model.

Costello et al. 2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be overly-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in Rosenberg et al. 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.

@@

tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
    rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
  select(`FAO Area`,`SOFIA % Overfished`, everything())

 tmp %>% 
  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

## Experiment Structure {.unnumbered}

We assessed the ability of Fisheries Management Index (FMI), swept area ratio (SAR), and effort data to improve estimates of the state of global unassessed fish populations. We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history. *B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data. This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries.

We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM. We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al. 2017 [@froese2017] (See Table.S1). We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished).

@@

We allow for process error *p* (in the manner of the stochastic stock reduction analysis (SRA) suggested by Walters *et al.* 2006 [@walters2006]). This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average. Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).

@@

All of our estimates are Bayesian in nature. sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent). Unless there is an abundance index to fit to the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. Any model results that are not based on an index of abundance are estimated through stock reduction analysis. All stock reduction analysis style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.

@@

<!-- We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history. Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero. -->

Suppose that the only thing we observe from a fishery is a catch history. Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past. In the absence of any data to fit to, the a stock reduction algorithm  (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure. This creates a problem for the Bayesian nature of our analysis. Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*. Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step. Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).

<!-- This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood.  -->

Borel's Paradox poses a particular problem  for the SRA version of sraplus. Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020). This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm.

We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000]. Our solution amounts to a two-step sample-importance-resampling (SIR) algorithm. We first run the standard SIR algorithm as described above. We then break the resulting draws into bins based on terminal stock status, and calculate the mean sampling probability of each bin. The net result of this is that it allows users to place explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior, rather than creating a complicated and biased prior on stock status based on a mixture of explicit and implicit priors. See Supplementary Information for a detailed explanation of this problem and our solution.

@@

Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus. We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database. We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types. Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries. For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on B/B~MSY~ or F/F~MSY~ in the most recent year of the fishery usable by sraplus (See Supplementary Information).

All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017]. The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus. Given the scope of this analysis, we do not claim that the regressions used here are the best possible model relating these data with the fishery status indicators of interest. Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.

@@

We performed a value-of-information (VOI) assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries. The VOI analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RAM, and comparing the estimated values to the values reported in RAM. We generate fits for 3000 combinations of a RAM stock and available data. For any one draw, we randomly sample a RAM stock and a list of available data and data quality. For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance. The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized.

@@

We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index scores, and swept area ratio (SAR) values (Table.\@ref(tab:data)).

@@
